\documentclass[a5paper]{book}
\usepackage[a5paper]{geometry}

\title{Praying through the Psalms}
\author{Jesse Lu}

\newcommand{\q}[1]{\textit{#1}}
\newcommand{\Q}[1]{\begin{quote}#1\end{quote}}
\begin{document}
\maketitle
\tableofcontents

\chapter{Psalm 1}
Psalm 1 is the first psalm in the book of Psalms
    because it forms the foundation of every believer's life.
Why is this psalm so critical to the Christian life?
Because it identifies what must be the distinguishing characteristic
    of every Christian---a delight for God's word (v. 3).

\section{The separation of the righteous and the ungodly}
The first truth we must understand from Psalm 1
    is that there are only two kinds of people on earth:
    the righteous and the ungodly.
Every human being is either a righteous or an ungodly person.
Every human is either in the category of the righteous or 
    else must be in the category of the ungodly;
    no one can be in both categories at the same time,
    and there is no third category that anyone can belong to.

\subsection{The general identification of the ungodly}
We know this to be true, first,
    because of the very general description of the ungodly in Psalm 1.
The ungodly are very simply described in verse 4 in this way, 
    \q{The ungodly are not so};
    which simply means that the ungodly man is not like the righteous man
    (who is described in verses 1-3)!

This means that the psalmist is not referring to some kind of extreme sinner
    when he uses the terms \emph{ungodly}, \emph{sinner}, or \emph{scoffer}.
Instead, the psalmist is simply referring to \emph{those who are not righteous!}
An equally valid translation for the word \q{ungodly}
    would be the word \q{unrighteous}.

This is why every human belongs to one and only one of these two groups
    and not to any third or fourth group---because 
    every human must either be righteous or else, must not.

\subsection{The divine separation of the ungodly}
The second reason why we can know that
    everyone is either righteous or ungodly is
    because these two groups are divinely separated by God's judgment.
Verse 5 reads,
    \Q{Therefore the ungodly shall not stand in the judgement, \\
    Nor sinners in the congregation of the righteous.}

A judgment results in only two possible outcomes:
    you either pass the judgment, or you don't.
Why is it impossible to be in both groups at the same time?
Because it is impossible to both pass and fail a judgement.
Instead, we will either stand in the judgement of God,
    or else we will not (verse 5a).

The second half of verse 5 further strengthens the notion
    of humanity divided into two distinct groups,
    because it refers to the congregation (or grouping) of the righteous
    which sinners are not a part of.
Therefore, everyone is either one of the righteous,
    who have a place in this congregation of the righteous,
    or a sinner, who does not.
    
\section{The distinguishing characteristic of the righteous}
The fact that men are only divided into either 
    the righteous or else the ungodly 
    then begs the questions,
    ``Who are the righteous?'' and
    ``How can I be part of the righteous?''
These questions are answered in verse 1 and 2,
    which expose the distinguishing characteristic of the righteous.
    \Q{Blessed is the man \\
    Who walks not in the counsel of the ungodly, \\
    Nor stands in the path of sinners, \\
    Now sits in the seat of the scornful; \\
    But his delight is in the law of the Lord, \\
    And in His law he meditates day and night.}
    
The distinguishing characteristic of the righteous man,
    in the negative sense,
    is that his life is antithetical to that of the ungodly.
Verse 1 says that he does not walk in their counsels, 
    stand in their paths, or sit in their seats.
This is not to say that the righteous is isolated and
    has no interaction with the ungodly;
    rather, it means that when the righteous
    come in contact with the ungodly,
    the ungodly man is meeting someone who's life is heading
    in a completely opposite direction from his own.

This contrast becomes clear as the central distinguishing characteristic
    of the righteous is presented in verse 2,
    ``But his delight is in the law of the Lord,''.

How is the righteous man identified? 
And how can we test if we are righteous men or women ourselves?
Psalm 1 gives us a clear test:
    Do we delight in the Scripture?

The test is \emph{not} whether we know the Bible,
    or simply read the Bible,
    or go to church to hear the Bible preached.
Psalm 1 does not even identify the righteous as the one
    who \emph{has delighted} in God's Word.
No, the righteous man of God is known by this primary characteristic:
    a joyous love and delighting in the word of God.

What is the root cause of trouble in the life of the believer?
And what is origin of our sin and unfaithfulness to our Lord?
Psalm 1 informs us that our root problem is our delighting in other things,
    over and above God's word.

And not only does Psalm 1 identify the root origin of our sin,
    but it also gives us the remedy for our situation:
    to delight in God's word.

The righteous man's delight in God's word is the overpowering characteristic
    of his life.
Without exaggeration, we may say that he is obsessed with Scripture.
Verse 2 ends, ``And in His law he meditates day and night.''
The faithful Christian life is characterized by
    an all-consuming love and joy in the word of God.
To what extent?
To the extent that it is his continual meditation throughout each day.
    
\section{The blessing of the righteous}
Having seen the \emph{separation} of the righteous,
    and the \emph{distinguishing characteristic} of the righteous,
    we now turn our attention to the \emph{blessing} of the righteous.
This is, in fact, the main point of Psalm 1 which begins with, 
    ``Blessed is the man\ldots''.
Verse 3 describes the blessing of the one who delights in the law of the Lord 
    in this way,
    \Q{He shall be like a tree \\
    Planted by the rivers of water, \\
    That brings forth its fruit in its season, \\
    Whose leaf also shall not wither, \\
    And whatever he does shall prosper.}

This verse describes the blessed life of the righteous man by way of analogy,
    ``He shall be like a tree'',
    and what a remarkably blessed tree the righteous man is!
First, we see that this tree is abundantly supplied
    in that it is, ``Planted by the rivers of water,''.
This means that the tree does not need to depend on the weather (i.e. rainfall)
    for its nourishment;
    instead, it is constantly nourished,
    not just by a river,
    but even a multitude of rivers!
What is the blessing of the righteous man?
The righteous man's blessing is to be abundantly and continually supplied for
    and provided for by God in every situation.

Jesus' words from John 15 describe a similar man
    as a branch who is connected to the vine that is Jesus Christ Himself.
This man, abiding in Jesus Christ, is not only perfectly provided for, 
    but is fruitful as well (John 15:5).
This fruitfulness is also seen in Psalm 1:3, 
    ``That brings forth its fruit in its season,''.
\end{document}
